Figure 6. Differential Protein expression

### 2022-02-24 reports/manuscript/Figure6.Rmd
Tweaking plot sizes and output

---
title: "Figure 6. Differential Protein expression"
output: 
  pdf_document:
    fig_caption: yes
header-includes: 
- \usepackage{float}
editor_options: 
  chunk_output_type: console
---

```{r setup}
knitr::opts_chunk$set(
  fig.path="./Figures/",
  fig.process = function(filename){
    new_filename <- stringr::str_remove(string = filename, 
                                        pattern = "-1")
    fs::file_move(path = filename, new_path = new_filename)
    ifelse(fs::file_exists(new_filename), new_filename, filename)
})
#setwd("/Users/vilkal/work/Brolidens_work/Projects/Gabriella_repo/reports/rmarkdown/manuscript")
```

```{r message=FALSE, warning=FALSE, include=FALSE}
##################
# LOAD LIBRARIES #
##################
suppressWarnings({suppressMessages({suppressPackageStartupMessages({
  library(tidyverse)
  library(edgeR)
  library(openxlsx)
  library(rstatix)
  #remotes::install_github("czarnewski/niceRplots",force=T)
  library(niceRplots)
})  })  })

#############
# LODA DATA #
#############
datasets_all_samples <- readRDS("../../../results/datasets_all_samples.RDS")
Protein_V3 <- read_csv("../../../data/2020_04_20_LAMIQ_visit_3_singlets.csv")
cytokine_data <- read.csv("../../../data/210125_Cytokines_visit3CVL.csv")
metadata <- read.csv2("../../../results/metadata_integration.csv",row.names = 1, stringsAsFactors = F)

@@

Protein <- Protein_V3 %>%
  left_join(select(metadata, PatID, ID), by="PatID") %>%
  select(ID, contains("HPA")) %>%
  filter(!is.na(.$ID)) %>%
  pivot_longer(., -ID, names_to = "Antibody") %>%
  pivot_wider(names_from = "ID", values_from = "value" ) %>%
  filter(!(Antibody %in% to_remove)) %>%
  column_to_rownames(var = "Antibody")

# Replace missing values with average in sexwork_months column
m <- round(sum(metadata$sexwork_months, na.rm = T)/nrow(metadata))
metadata <- metadata %>%
  mutate(sexwork_months = ifelse(is.na(.$sexwork_months), m, .$sexwork_months))

@@

    # Study group annotation
    end <- table(pal[factor(names(samples))[ order(df[1])]])
    end <- end[order(factor(names(end), levels = pal))]
    end <- map_dbl(cumsum(end), ~ (.x * (par("usr")[2])/length(samples)-.017) )
    start <- c(0, map_dbl(end[1:4], ~.x+0.015))

    axis(1, at = c(start, end), label = F, pos =1.03 , xpd=T, col="white",col.ticks="black")
    map2(start, end, ~lines(x=c(.y, .x), y= c(1.03, 1.03), xpd=T, cex=.8))
    # text labels
    t <- map2(start, end, ~(.x+.y)/2)
    map2(t,levels(factor(names(samples))), ~text(x=.x, y=1.04, .y,xpd=T, cex=.8))

    # Boxes 
    if(gr == "Luminal_gr") {
      lines(x=c(0,.603,.603,0,0),
            y=c(.488,.488,1.006, 1.006,.488), col="red", lty=1, lwd = 1.5)
      lines(x=c(.39,.94,.94,.39,.39),
            y=c(.488,.488,-0.008, -0.008,.488), col="red", lty=1, lwd = 1.5)}
    if(gr == "Tissue_gr") {
      lines(x=c(.653,.95,.95,.653,.653),
            y=c(.735,.735,1.011, 1.011,.735), col="red", lty=1, lwd = 1.5)
      lines(x=c(.653,.95,.95,.653,.653),
            y=c(.735,.735,-0.012, -0.012,.735), col="red", lty=1, lwd = 1.5)
    }

    ## Meta info bars
    points( seq(0,1,length.out = length(samples)),rep(par("usr")[4],length(samples))+.005,
            col=pal[factor(names(samples))[ order(df[1])]],
            pch=15,xpd=T,cex=.7)

    points( seq(0,1,length.out = ncol(top_PROT_mfi)),rep(par("usr")[3],ncol(top_PROT_mfi))-.008,
            col=c("tomato","orange","#d7d7d7")[ factor(metadata[samples,"BV_Diagnosis_v3"][order(df[1])]) ],
            pch=15,xpd=T,cex=.5) 
    points( seq(0,1,length.out = ncol(top_PROT_mfi)),rep(par("usr")[3],ncol(top_PROT_mfi))-.022,
            col= c("#d7d7d7","tomato")[ factor(metadata[samples,"HIVstatus"][order(df[1])]) ],
            pch=15,xpd=T,cex=.5)

@@

imap(p_list, ~write.xlsx(.x,file=paste0("./Suppl.Tbl/","Suppl.Tbl.",.y,"_Proteins.xlsx"))) # Excel
write.xlsx(list("L2T2 vs L2T3"=dpe_LT$top),file=paste0("./Suppl.Tbl/","Suppl.Tbl.17"," L2T2_L2T3_Proteins.xlsx"))
```

@@

**Figure 3. Protein Heatmap**. **a** and **b** top differentially expressed proteins scaled by row, for Luminal and Tissue respectively. (p-value < 0.05) **b** and **d** sig. cytokines for Luminal and Tissue respectively.

>>>>

---
title: "Figure 6. Differential Protein expression"
geometry: "left=2cm,right=2cm,top=2cm,bottom=2cm"
output: 
  pdf_document:
    fig_caption: yes
header-includes: 
- \usepackage{float}
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  fig.width = 6.6929133858,
  fig.path="./Figures/",
  dev=c('pdf', 'png'), dpi=300,
  fig.process = function(filename){
    new_filename <- stringr::str_remove(string = filename, 
                                        pattern = "-1")
    fs::file_move(path = filename, new_path = new_filename)
    ifelse(fs::file_exists(new_filename), new_filename, filename)
})
# setwd("/Users/vilkal/work/Brolidens_work/Projects/broliden_5325/reports/manuscript")
```

```{r message=FALSE, warning=FALSE, include=FALSE}
##################
# LOAD LIBRARIES #
##################
suppressWarnings({suppressMessages({suppressPackageStartupMessages({
  library(tidyverse)
  library(edgeR)
  library(openxlsx)
  library(readxl)
  library(rstatix)
  #remotes::install_github("czarnewski/niceRplots",force=T)
  library(niceRplots)
})  })  })

#########
# PATHS #
#########
result_dir <- "./Suppl.Tbl/"
if( isFALSE(dir.exists(result_dir)) ) { dir.create(result_dir,recursive = TRUE) }

#############
# LODA DATA #
#############
datasets_all_samples <- readRDS("../../results/03_normalize_data_output/datasets_all_samples.RDS")
Protein_V3 <- read_xlsx("../../data/Protein_norm_MFI_Bradley_et.al.xlsx")
cytokine_data <- read.csv("../../data/210125_Cytokines_visit3CVL.csv")
metadata <- read.csv("../../data/metadata.csv",row.names = 1, stringsAsFactors = F)

@@

Protein <- Protein_V3 %>%
  select(ID, contains("HPA")) %>%
  filter(ID %in% metadata$ID) %>%
  pivot_longer(., -ID, names_to = "Antibody") %>%
  pivot_wider(names_from = "ID", values_from = "value" ) %>%
  filter(!(Antibody %in% to_remove)) %>%
  column_to_rownames(var = "Antibody")

@@

    # Study group annotation
    end <- table(pal[factor(names(samples))[ order(df[1])]])
    end <- end[order(factor(names(end), levels = pal))]
    end <- map_dbl(cumsum(end), ~ (.x * (par("usr")[2])/length(samples)-.017) )
    start <- c(0, map_dbl(end[1:4], ~.x+0.015))
    end <- c(end[1:length(end)-1], par("usr")[2]-.005)

    axis(1, at = c(start, end), label = F, pos =1.032 , xpd=T, col="white",col.ticks="black")
    map2(start, end, ~lines(x=c(.y, .x), y= c(1.032, 1.032), xpd=T, cex=.8))
    # text labels
    t <- map2(start, end, ~(.x+.y)/2)
    map2(t,levels(factor(names(samples))), ~text(x=.x, y=1.043, .y,xpd=T, cex=.8))

    # Boxes 
    if(gr == "Luminal_gr") {
      lines(x=c(0,.603,.603,-.002,-.002),
            y=c(.488,.488,1.008, 1.008,.488), col="red", lty=1, lwd = 1.5)
      lines(x=c(.39,.94,.94,.39,.39),
            y=c(.488,.488,-0.008, -0.008,.488), col="red", lty=1, lwd = 1.5)}
    if(gr == "Tissue_gr") {
      lines(x=c(.653,.95,.95,.653,.653),
            y=c(.735,.735,1.011, 1.011,.735), col="red", lty=1, lwd = 1.5)
      lines(x=c(.653,.95,.95,.653,.653),
            y=c(.735,.735,-0.012, -0.012,.735), col="red", lty=1, lwd = 1.5)
    }

    ## Meta info bars
    if(gr == "Luminal_gr"){size = .5}else{size=.6}
    points( seq(0,1,length.out = length(samples)),rep(par("usr")[4],length(samples))+.005,
            col=pal[factor(names(samples))[ order(df[1])]],
            pch=15,xpd=T,cex=.7)

    points( seq(0,1,length.out = ncol(top_PROT_mfi)),rep(par("usr")[3],ncol(top_PROT_mfi))-.008,
            col=c("tomato","orange","#d7d7d7")[ factor(metadata[samples,"BV_Diagnosis_v3"][order(df[1])]) ],
            pch=15,xpd=T,cex=size) 
    points( seq(0,1,length.out = ncol(top_PROT_mfi)),rep(par("usr")[3],ncol(top_PROT_mfi))-.022,
            col= c("#d7d7d7","tomato")[ factor(metadata[samples,"HIVstatus"][order(df[1])]) ],
            pch=15,xpd=T,cex=size)

@@

imap(p_list, ~write.xlsx(.x,file=paste0(result_dir,"Suppl.Tbl.",.y,"_Proteins.xlsx"))) # Excel
write.xlsx(list("L2T2 vs L2T3"=dpe_LT$top),file=paste0(result_dir,"Suppl.Tbl.17"," L2T2_L2T3_Proteins.xlsx"))
```

@@

**Figure 6. Protein Heatmap**. **a** and **c** top differentially expressed proteins scaled by row, for Luminal and Tissue respectively. (p-value < 0.05) **b** and **d** sig. cytokines for Luminal and Tissue respectively.
